---
title: "University Students' Drinking Data "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(stringr)
library(forcats)

```

Cleaning data: 
- Renaming variables to make them more workable
- Separating drinks column into minDrinks and maxDrinks and making those variables     numeric


```{r }
drinksDS <- read_csv("Stats survey.csv")

drinksDS %>% 
  variable.names()
drinksDS_clean <- drinksDS %>% 
  rename(gender = "Your Sex?", 
         collegeYear2023 = "What year were you in last year (2023) ?", 
         grade12GPA = "Your Matric (grade 12) Average/ GPA (in %)", 
         major = "What faculty does your degree fall under?", 
         "2023GPA" = "Your 2023 academic year average/GPA in % (Ignore if you are 2024 1st year student)", 
         accomodationStatus2023 = "Your Accommodation Status Last Year (2023)", 
         monthlyAllowance2023 = "Monthly Allowance in 2023", 
         scholarship2023 = "Were you on scholarship/bursary in 2023?", 
         studyTime = "Additional amount of studying (in hrs) per week", 
         partyTime = "How often do you go out partying/socialising during the week?", 
         drinks = "On a night out, how many alcoholic drinks do you consume?", 
         missedClasses = "How many classes do you miss per week due to alcohol reasons, (i.e: being hungover or too tired?)", 
         failedModules = "How many modules have you failed thus far into your studies?", 
         relationshipStatus = "Are you currently in a romantic relationship?", 
         parentApproval = "Do your parents approve alcohol consumption?", 
         parentRelationship = "How strong is your relationship with your parent/s?") %>% 
  separate(drinks, c("minDrinks", "maxDrinks"), sep = "-", remove = FALSE, convert = TRUE) %>% 
  mutate(minDrinks = as.numeric(str_remove(minDrinks, "\\+"))) %>% 
  select(-Timestamp) %>% 
  as_tibble()

```
Quick glimpse of data: Variables, Datetypes, descriptive statistics....

```{r}
drinksDS_clean %>% 
  summary()
```

Counts of majors: 
This data is heavily skewered towards Economic & Management Sciences majors

```{r}
drinksDS_clean %>% 
  count(major)
```
Counts of gender (Male/Female): 
Data collected from mostly males

```{r}
drinksDS_clean %>%
  count(gender)
```


Counts of grade level in 2023

```{r}
drinksDS_clean %>% 
  count(collegeYear2023)

```
```{r}
drinksDS_clean %>% 
  count(accomodationStatus2023)
```
Count of how many times students go out per week
Students seem to go out mostly on weekends

```{r}
drinksDS_clean %>% 
  count(partyTime, sort = TRUE)
```

Count of Students' monthly allowance

```{r}

drinksDS_clean %>% 
  count(monthlyAllowance2023)

```

Count of Students' relationship status

```{r}
drinksDS_clean %>% 
  count(relationshipStatus)
```

```{r}
drinksDS_clean %>% 
  count(failedModules)
```
Parent Approval

```{r}
drinksDS_clean %>% 
  count(parentApproval)
```
Distribution of GPA in Senior Year of high school

```{r}
drinksDS_clean %>% 
  group_by(grade12GPA) %>% 
  summarise(count = n()) %>% 
  ggplot() + 
  geom_point(aes(grade12GPA, count)) + 
  labs(x = "Senior Year GPA in %") 
```
Distribution of GPA for 2023 college year

```{r}
drinksDS_clean %>% 
  count(`2023GPA`) %>% 
  ggplot() + 
  geom_point(aes(`2023GPA`, n)) + 
  labs(x = "Senior Year GPA in %")
```
Run Chi Sq test to see if we have a difference between Males and Females. 
We cannot reject the null hypothesis

```{r}
drinksDS_clean %>%
  count(gender)

studentGender <- drinksDS_clean %>%
  select(gender) %>% 
  filter(gender %in% c("Female", 
                       "Male")) %>% 
  mutate(gender = fct_drop(gender))

studentGender_table <- table(studentGender)
View(studentGender_table) 

chisq.test(studentGender_table)
```
How many drinks do different gender have when going out

```{r}
drinksDS_clean %>% 
  group_by(gender) %>% 
  filter(gender != is.na(gender)) %>% 
  count(gender, minDrinks) %>% 
  ggplot(aes(minDrinks, n, color = gender)) +
  geom_line() +
  labs(x = "Minimum amount of drinks when going out", 
       y = "Count", 
       title = "Amount of drinks university students have", caption = "Data Source:https://www.kaggle.com/datasets/joshuanaude/effects-of-alcohol-on-student-performance") + 
  theme(plot.caption = element_text(size = 5, hjust = -.25))+
  scale_color_brewer(palette = "Dark2")
```


How many drinks do different college year students have when going out
#Removed Missing Data
```{r}
drinksDS_clean %>% 
  select(collegeYear2023, `2023GPA`, major, grade12GPA, studyTime, partyTime, minDrinks) %>% 
  group_by(collegeYear2023, minDrinks) %>% 
  filter(collegeYear2023 != is.na(collegeYear2023)) %>% 
  summarise(count = n()) %>% 
  ggplot() + 
  geom_area(aes(minDrinks, count, fill = collegeYear2023)) + 
  theme_classic() +
  labs(x = "Minimum Amount of Drinks When Going Out", fill = "2023 College Year", title = "How many alcoholic drinks do college students have", caption = "Data Source:https://www.kaggle.com/datasets/joshuanaude/effects-of-alcohol-on-student-performance") + 
  theme(plot.caption = element_text(size = 5, hjust = -.25)) +
  scale_fill_brewer(palette = "Spectral")

```

How many times a week do different major go out

```{r}
drinksDS_clean %>%
  group_by(partyTime, major) %>%
  filter(major != is.na(major)) %>% 
  count(partyTime, major, sort = TRUE) %>% 
  ggplot(aes(fct_reorder(partyTime, n, .fun = sum, .desc = TRUE), n, fill = major)) + 
  geom_col() + 
  labs(x = "# of times student go out", 
       y = "Amount of students that go out", 
       fill = "Major/Focus",
  caption = "Data Source:https://www.kaggle.com/datasets/joshuanaude/effects-of-alcohol-on-student-performance") + 
  theme(plot.caption = element_text(size = 5, hjust = -.25),
        axis.text.x = element_text(angle = 45, size = 5, vjust = .5))+
  scale_fill_brewer(palette = "Spectral")

  
```

How many drinks do different majors have when going out

```{r}
drinksDS_clean %>% 
  filter(partyTime == "Only weekends", 
         major != is.na(major)) %>%  
  count(major, minDrinks) %>% 
  mutate(minDrinks = reorder(minDrinks, n, FUN = sum)) %>% 
  ggplot(aes(minDrinks, n, fill = major)) + 
  geom_col(aes()) + 
  labs(title = "How many drinks do different major have on the weekends", 
       x = "Minimum amount of drinks", 
       y = "count", 
       caption = "Data Source:https://www.kaggle.com/datasets/joshuanaude/effects-of-alcohol-on-student-performance")+
  theme(plot.caption = element_text(size = 5, hjust = -.25),
        axis.text.x = element_text(angle = 45, size = 5, vjust = .5))+
  scale_fill_brewer(palette = "Spectral")

```




